---
title: "PA1_template.Rmd"
author: "Francis Ivan Clado"
date: "May 10, 2019"
output: html_document
---

## Loading and preprocessing the data
```{r}
library(ggplot2)
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Use the tapply function to calculate the total steps per day.
```{r}
total <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
```

Load the ggplot2 package and use the qplot function to create a histogram.
```{r}
library(ggplot2)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
```

Use the mean and median function.
```{r}
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```


## What is the average daily activity pattern?
Call the aggregate function to get the average steps for each 5-minute interval.
```{r}
intervalsteps <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
```

Use ggplot to create a time series plot.
```{r}
ggplot(data=intervalsteps, aes(x=interval, y=steps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps taken")
```

Use the which and max function to determine Which 5-minute interval contains the maximum number of steps.
```{r}
intervalsteps[which.max(averages$steps),]
```


## Imputing missing values
Create a logical vector that indicates whether a value in data$steps is null, then use the length function to count the null values. Replace the null values with the mean of the number of steps.
```{r}
nas <- is.na(data$steps)
length(nas[nas==TRUE])
data[is.na(data)] <- mean(data$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
Create a function that determines if a given date is a weekday or weekend.
```{r}
daytype <- function(date) {
type <- ""  
type <- ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
type 
}
```

Convert the values in data$date into a date class. Use sapply to use the created function on each date. Convert the daytype values into a factor variable with two levels.
```{r}
data$date <- as.Date(data$date)
data$daytype <- sapply(data$date, FUN=daytype)
data$daytype <- as.factor(data$daytype)
```

Use the aggregate function and ggplot to create a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
panelplot <- aggregate(steps ~ interval + daytype, data=data, mean)
ggplot(panelplot, aes(interval, steps)) + geom_line() + facet_grid(daytype ~ .) +
  xlab("5-minute interval") + ylab("Number of steps")
```

